---
output: github_document
references:
- id: analyiticalSIR
  title: Analytical solution of the SIR-model for the temporal evolution of epidemics. Part A time-independent reproduction factor
  author:
  - family: Kröger
    given: M
  - family: Schlickeiser
    given: R
  container-title: Journal of Physics A Mathematical and Theoretical
  volume: 53
  URL: 'https://iopscience.iop.org/article/10.1088/1751-8121/abc65d'
  DOI: 10.1088/abc65d
  publisher: IOP Science
  type: article-journal
  issued:
    year: 2020
    month: 12
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# excalibur

<!-- badges: start -->
[![R-CMD-check](https://github.com/mrc-ide/excalibur/actions/workflows/R-CMD-check.yaml/badge.svg?branch=main)](https://github.com/mrc-ide/excalibur/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/mrc-ide/excalibur/actions/workflows/test-coverage.yaml/badge.svg?branch=main)](https://github.com/mrc-ide/excalibur/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

**Ex**pediating **Calib**rations in yo**ur** epidemic models

The goal of excalibur is to explore the necessity of calibrations on ODE 
epidemic models that are intended to predict future epidemic trajectories.

* :warning: This package is work in progress

## Installation

You can install the released version of *excalibur* from [its Github repository](https://github.com/mrc-ide/excalibur) with:

``` r
devtools::install_github("mrc-ide/excalibur")
```

This package requires *Odin*. *Deriv* and *stringr* need to be installed if you
wish to calculate your own derivatives if using the derivative approximation 
method.

## SIRD Model

### Examples

Estimating the current state for a simple SIRD model:

```{r SIRD example1}
library(excalibur)
##Simulate some deaths to demonstrate calculating the current state of the model
#Set up an SIRD model
model <- setSIRD(N = 100, Beta = 1, Gamma = 1/3, ProbOfDeath = 0.1, I0 = 1)
#Generate some results at time 10 from the start of the epidemic
time <- 10
deaths <- simulate(model, t = time)$D
##Calculate the current state of the model
model <- calculateCurrentState(model, time, deaths)
#print the current state
print(currentState(model))

##Limitations
#Currently I is calculated by subtracting from N, hence will not be exact due to
#rounding in storage etc.
print(simulate(model, t=time))
```

Estimating the current state for a simple SIRD model, with time-varying Beta:

```{r SIRD example2}
library(excalibur)
#Set up an SIRD model, with constant values of Beta that change at the times given
model <- setSIRD(N = 100, Beta = c(5,0.1,2), Gamma = 1/3, ProbOfDeath = 0.1, 
                I0 = 1, changeTimes = c(4,6))
#Generate some results, since this is time-varying we need these at the end of
#each of the changeTimes as well as at the current time
time <- 10
deaths <- simulate(model, t = c(4,6,time))$D
##Calculate the current state of the model
model <- calculateCurrentState(model, time, deaths)
#print the current state
print(currentState(model))
#for comparison
print(simulate(model, t=time))
```

### Method
Consider the SIRD model with a known number of deaths $D(t)$ a particular time $t$, where $\beta$ is the transmission parameter, $\gamma$ the rate of recovery, and $\alpha$ the rate of death (calculated from the probability of death).
Then the number of people in the recovered population at time $t$ is, 

$R(t)=(D(t)-D(0))\times\gamma/\alpha + R(0).$

To simplify, let $D(i)+R(i)=\mathrm{\textit{DR}}(i)$.
Then by a similar argument to that laid out in [@analyiticalSIR], for a constant value of $\beta$,

$S(t) = S(0)\times \exp((\mathrm{\textit{DR}}(0)-\mathrm{\textit{DR}}(t))\times\beta/(N\times(\alpha+\gamma))).$

Then expanded to the time-varying $\beta$ scenario,

$S(t)=S(0)\times\exp((\mathrm{sum}^n_{i=0}(\mathrm{\textit{DR}}(t_{c,i})-\mathrm{\textit{DR}}(t_{c,i+1}))\times\beta_i)/(N\times(\alpha + \gamma)))$

where $t_{c,i} \leq t < t_{c,i+1}$ means that $\beta(t) = \beta_i$ and $t_{c,0} = 0$.

## SEIRD Model

### Examples

Estimating the current state for a simple SEIRD model, with time-varying Beta:

```{r SEIRD example}
library(excalibur)
#Set up an SIRD model, with constant values of Beta that change at the times given
model <- setSEIRD(N = 100, Beta = c(5,0.1,2), Gamma = 1/3, Lambda = 1/2, ProbOfDeath = 0.1, 
                I0 = 1, changeTimes = c(4,6))
#Generate some results, since this is time-varying we need these at the end of
#each of the changeTimes as well as at the current time
time <- 10
deaths <- simulate(model, t = c(4,6,time))$D
##Calculate the current state of the model
model <- calculateCurrentState(model, time, deaths, nderiv = 7)
#print the current state
print(currentState(model))
#for comparison
print(simulate(model, t=time))

#if we want to check the optimization we can also plot the derivative
model <- calculateCurrentState(model, time, deaths, nderiv = 7, plotDeriv = TRUE)
```

### Method
This method assumes that the n-th derivative of D is 0 and solves the resulting 
equation for I. n is determined by the argument nderiv and the derivative has been
calculated symbolically with the package *Deriv*. The higher n is the closer this
approximation gets to the real solution, however for certain set of parameters,
increase n also flattens the derivative, making it harder to optimise for I.
